---
title: "ISLR Chapter 3 Lab 3: Boston Q & A Part 1"
date: "January 30, 2020"
output: html_document
---

```{r setup, echo = FALSE}
require("MASS")
require("ISLR")
require("tidyverse")
require("GGally")
require("broom")
```

# A regression analysis of Boston Housing data

Consider the dataset `Boston`.  How many rows and columns are there?  What are the variables and their types?

```{r}
Boston <- as_tibble(Boston)
###Enter code here to answer the questions above.

summary(Boston)
str(Boston)
class(Boston)

```

In this dataset we are interested in finding whether there is a relationship between `medv` and some of the input variables, how strong the relationship is, whether a linear relationship is a good fit or not, what inferences can be made, how well can we predict values, etc.

But first, "Plot the data."  Make pairwise scatterplots of `medv`, `nox`, `rm`, and `lstat` using `ggpairs`.  Observations?

```{r}

# HW plot only 4 columns
Boston %>% 
  ggpairs(cols = c(5, 6, 13, 14))
```

## 1. Is there a relationship between the level of nitrous oxide (nox), number of rooms of a house (rm), or the economic status of a neighborhood (lstat) on the median value of the homes (medv)?

Our first goal should be to determine whether the data provide evidence of an association between the input variables $nox$, $rm$, and $lstat$ and the output variable $medv$.


### The $F$-test

This question can be answered by using an $F$-test.
Here are the five steps of the overall $F$-test for regression [1](http://facweb.cs.depaul.edu/sjost/csc423/documents/f-test-reg.htm)
 
1. State the null and alternative hypotheses in the context of the problem.
 
  $H_0:   ??_1 = ??_2 = ... = ??_{p-1} = 0$
 
  $H_a:   ??_j \neq 0$, for at least one value of $j$
 
2. Compute the test statistic assuming that the null hypothesis is true.
 
  This is done for use by R, but the calculation is 
  
  $ F = \frac{\textrm{explained variance}}{\textrm{unexplained variance}}.$
 
3. Calculate the $p$-value for the $F$-statistic for this model in an $F$ distribution with $(p - 1 , n - p + 1)$.  Again, R does this for us.  Reject the null hypothesis if the $p$-value is less than a predetermined value of $\alpha$.  For the purposes of this class, we will use $\alpha = 0.01$, but that would be determined by your organization or by the end user.
 
### Applying the $F$-test

This question can be answered by fitting a multiple regression model
of $medv$ onto $nox$, $rm$, and $lstat$ and testing the null hypothesis that in the equation 

$$medv = \beta_{nox} x_{nox} + \beta_{rm} x_{rm} + \beta_{lstat} x_{lstat}$$ 
all the coefficients are zero.  That is, the null hypothesis is 

$$H_0 : \beta_{nox} = \beta_{rm} = \beta_{lstat} = 0.$$
If we reject the null hypothesis, then there is evidence that at least one of the $\beta_k$ is not zero.

```{r}
# create the model object, output the F-statistic
#medv.fit <- lm(....)
medv.fit <- lm(medv ~ nox + rm + lstat, data = Boston)
glance(medv.fit)

```

The $p = ...$ for the $F$-statistic in this multivariable linear regression model is ..... than $0.01$, indicating that we would .... the null hypothesis that the coefficients are all 0.
Thus,  there .... evidence that at least one of the coefficients is not 0.

The conclusion to the $F$-test  is that there is or is not evidence of  a relationship, but testing the $F$-statistic does not indicate whether the relationship is strong or not, it does not indicate whether a linear fit is the right choice, and it does not indicate which or how many of the $\beta_k$ is non-zero.

## 2. How strong is the relationship between the input variables `nox`, `rm`, and `lstat` and the median value of the homes?

Can we predict the median value of a home with
a high level of accuracy using these input variables? Or is
a prediction of `medv` based on these three input factors only slightly
better than a random guess? 

There are two measures of model model accuracy in ISLR: percent error in predictions and the adjusted-$R^2$.

First,
the $RSE$ estimates the standard deviation of the response from the
population regression line. For this model, the RSE is .....
units while the mean value for the response is ...., indicating a
percentage error of roughly  25%, which is a pretty big error and the end users would have to decide if that's okay.... . (See ISLR Section 3.1.3)

Second, the adjusted $R^2$ statistic records
the percentage of variability in the response that is explained by
the predictors. The predictors explain almost ..... of the variance in
median house value.

Based on these two values, the percent error and the adjusted $R^2$, we could say that this is a moderately strong model.

```{r}
# Write code that will calculate the RSE and adjusted-R^2
y_bar <- mean(Boston$medv)

```

In conclusion, we can predict median house values to within .... of the average median house value and the model explains ..... of the variance in median house value.  Thus, this is a (strong, weak) model.


#### 3. Which of `nox`, `rm`, and `lstat` contribute to the median value of the homes?

Do all three input variables chosen contribute to the median value of the homes,
or do just one or two of the input variables  contribute? 

To answer this question, we can examine the $p$-values associated with
each predictor's $t$-statistic (see ISLR Section 3.1.2). In this model, the $p$-value(s) for 'rm' and 'stat', esscentially zero,  ....  is/are low,
but the $p$-value(s) for ....  is/ar not. This suggests that only .... and
is/are related to sales. (In ISLR Chapter 6 we explore this question in
greater detail.)

```{r}
# Show the statistics associated with the coefficients
tidy(medv.fit)
```


## 4. How accurately can we estimate the effect of each of 'nox', 'rm' and 'lstat' on median house value?

This question is asking how confident are we in the coefficients of the model.  For every unit change in each of `nox`, `rm`, and `lstat`, by what amount will the median value of homes increase or decrease? How accurately can we predict this
amount of increase?

ISLR Section 3.1.2 shows that the standard error of $\hat{??}_j$ can be used
to construct confidence intervals for $??_j$. The the 95 % confidence intervals are as follows: 

* (-1.85 - 2(2.65), -1.85r + 2(3.65)) for `nox`,

* (5.12 - 2(0.447), 5.12 + 2(0.447)) for `rm`, and 

* (-.062 - 2(0.052), -.062 + 2(0.052)) for `lstat`. 

The confidence interval(s) for ........ is/are narrow and far from zero, providing evidence that ..... are related to `medv`.  But the interval(s)
for .... include(s) zero, indicating that there is a chance that the the actual coefficient is zero. 

In the next lab we will continue to look at the analysis of this model.

##  Crime rates

Choose 2  variables from the Boston data set that might predict the crime rate and 1 variable that would not predict the crime rate.  Reproduce the analysis above.

In this dataset we are interested in finding whether there is a relationship between `medv` and some of the input variables, how strong the relationship is, whether a linear relationship is a good fit or not, what inferences can be made, how well can we predict values, etc.

But first, "Plot the data."  Make pairwise scatterplots of `medv`, `nox`, `rm`, and `lstat` using `ggpairs`.  Observations?

```{r}

# HW plot only 4 columns
Boston %>% 
  ggpairs(columns = c("crim","tax", "dis", "age"))
```

## 1. Is there a relationship between the level of nitrous oxide (nox), number of rooms of a house (rm), or the economic status of a neighborhood (lstat) on the median value of the homes (medv)?

Our first goal should be to determine whether the data provide evidence of an association between the input variables $nox$, $rm$, and $lstat$ and the output variable $medv$.


### The $F$-test

This question can be answered by using an $F$-test.
Here are the five steps of the overall $F$-test for regression [1](http://facweb.cs.depaul.edu/sjost/csc423/documents/f-test-reg.htm)
 
1. State the null and alternative hypotheses in the context of the problem.
 
  $H_0:   ??_1 = ??_2 = ... = ??_{p-1} = 0$
 
  $H_a:   ??_j \neq 0$, for at least one value of $j$
 
2. Compute the test statistic assuming that the null hypothesis is true.
 
  This is done for use by R, but the calculation is 
  
  $ F = \frac{\textrm{explained variance}}{\textrm{unexplained variance}}.$
 
3. Calculate the $p$-value for the $F$-statistic for this model in an $F$ distribution with $(p - 1 , n - p + 1)$.  Again, R does this for us.  Reject the null hypothesis if the $p$-value is less than a predetermined value of $\alpha$.  For the purposes of this class, we will use $\alpha = 0.01$, but that would be determined by your organization or by the end user.
 
### Applying the $F$-test

This question can be answered by fitting a multiple regression model
of $medv$ onto $nox$, $rm$, and $lstat$ and testing the null hypothesis that in the equation 

$$medv = \beta_{nox} x_{nox} + \beta_{rm} x_{rm} + \beta_{lstat} x_{lstat}$$ 
all the coefficients are zero.  That is, the null hypothesis is 

$$H_0 : \beta_{nox} = \beta_{rm} = \beta_{lstat} = 0.$$
If we reject the null hypothesis, then there is evidence that at least one of the $\beta_k$ is not zero.

```{r}
# create the model object, output the F-statistic
#medv.fit <- lm(....)
medv.fit <- lm(medv ~ nox + rm + lstat, data = Boston)
glance(medv.fit)

```

The $p = ...$ for the $F$-statistic in this multivariable linear regression model is ..... than $0.01$, indicating that we would .... the null hypothesis that the coefficients are all 0.
Thus,  there .... evidence that at least one of the coefficients is not 0.

The conclusion to the $F$-test  is that there is or is not evidence of  a relationship, but testing the $F$-statistic does not indicate whether the relationship is strong or not, it does not indicate whether a linear fit is the right choice, and it does not indicate which or how many of the $\beta_k$ is non-zero.

## 2. How strong is the relationship between the input variables `nox`, `rm`, and `lstat` and the median value of the homes?

Can we predict the median value of a home with
a high level of accuracy using these input variables? Or is
a prediction of `medv` based on these three input factors only slightly
better than a random guess? 

There are two measures of model model accuracy in ISLR: percent error in predictions and the adjusted-$R^2$.

First,
the $RSE$ estimates the standard deviation of the response from the
population regression line. For this model, the RSE is .....
units while the mean value for the response is ...., indicating a
percentage error of roughly  25%, which is a pretty big error and the end users would have to decide if that's okay.... . (See ISLR Section 3.1.3)

Second, the adjusted $R^2$ statistic records
the percentage of variability in the response that is explained by
the predictors. The predictors explain almost ..... of the variance in
median house value.

Based on these two values, the percent error and the adjusted $R^2$, we could say that this is a moderately strong model.

```{r}
# Write code that will calculate the RSE and adjusted-R^2
y_bar <- mean(Boston$medv)

```

In conclusion, we can predict median house values to within .... of the average median house value and the model explains ..... of the variance in median house value.  Thus, this is a (strong, weak) model.


#### 3. Which of `nox`, `rm`, and `lstat` contribute to the median value of the homes?

Do all three input variables chosen contribute to the median value of the homes,
or do just one or two of the input variables  contribute? 

To answer this question, we can examine the $p$-values associated with
each predictor's $t$-statistic (see ISLR Section 3.1.2). In this model, the $p$-value(s) for 'rm' and 'stat', esscentially zero,  ....  is/are low,
but the $p$-value(s) for ....  is/ar not. This suggests that only .... and
is/are related to sales. (In ISLR Chapter 6 we explore this question in
greater detail.)

```{r}
# Show the statistics associated with the coefficients
tidy(medv.fit)
```


## 4. How accurately can we estimate the effect of each of 'nox', 'rm' and 'lstat' on median house value?

This question is asking how confident are we in the coefficients of the model.  For every unit change in each of `nox`, `rm`, and `lstat`, by what amount will the median value of homes increase or decrease? How accurately can we predict this
amount of increase?

ISLR Section 3.1.2 shows that the standard error of $\hat{??}_j$ can be used
to construct confidence intervals for $??_j$. The the 95 % confidence intervals are as follows: 

* (-1.85 - 2(2.65), -1.85r + 2(3.65)) for `nox`,

* (5.12 - 2(0.447), 5.12 + 2(0.447)) for `rm`, and 

* (-.062 - 2(0.052), -.062 + 2(0.052)) for `lstat`. 

The confidence interval(s) for ........ is/are narrow and far from zero, providing evidence that ..... are related to `medv`.  But the interval(s)
for .... include(s) zero, indicating that there is a chance that the the actual coefficient is zero. 

In the next lab we will continue to look at the analysis of this model.

