---
title: "3.3.1 Qualitative Predictors "
author: "Josie Clarke & Daniel Camacho"
date: "February 14th, 2020"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
require(ISLR)
require(tidyverse)
require(broom)
```

## Including Qualitative Variables into a Regression Model

Qualitative or factor variables as predictors:

- Gender
- Political affiliation
- Student status

They can be included in the model by creating an indicator or dummy variable.

Let's study the following scenarios:

- Predictors with only two levels
- Predictors with more than two levels


## Predictors with Only Two Levels

Consider the *Credit* data set

+ The aim here is to predict which customers will default on their credit card debt.

Based on the *gender* variable:

$$ x_{i1} =
  \begin{cases}
                                   1 & \text{if the $i$th person is female} \\
                                   0 & \text{if the $i$th person is male}
  \end{cases}$$

Regression equation:
$$\begin{aligned}
    y_{i} = {\beta_{0}} + {\beta_{1}x_{i1}} +  {\epsilon_i}
          =  \begin{cases}
                    {\beta_{0}} + {\beta_{1}} + {\epsilon_i} & \text{if the $i$th person is female} \\
                    {\beta_{0}}  + {\epsilon_i} & \text{if the $i$th person is male}\\
              \end{cases}
\end{aligned}$$

## Least squares coefficient estimates
Gender is enconded as a dummy variable in the model (Table 3.7)
```{r, echo = FALSE}
model.fit <- lm(Balance ~ Gender, 
               data = Credit)
tidy(model.fit)
# contrasts(Credit$Gender)
```


Average credit card debt:

- Males: $\$509.80$
- Females: $\$509.80 + \$19.73 = \$529.53$

p-value for the dummy variable is very high

## Arbitrary Coding Scheme
 The main difference is the way in which the coefficients are interpreted
$$ x_{i1} =
  \begin{cases}
                                   1 & \text{if the $i$th person is female} \\
                                   0 & \text{if the $i$th person is male}
  \end{cases}$$
  
$$ x_{i1} =
  \begin{cases}
                                   1 & \text{if the $i$th person is female} \\
                                   -1 & \text{if the $i$th person is male}
  \end{cases}$$
  
Regression equation:
$$\begin{aligned}
    y_{i} = {\beta_{0}} + {\beta_{1}x_{i1}} +  {\epsilon_i}
          =  \begin{cases}
                    {\beta_{0}} + {\beta_{1}} + {\epsilon_i} & \text{if the $i$th person is female} \\
                    {\beta_{0}} - {\beta_{1}} + {\epsilon_i} & \text{if the $i$th person is male}\\
              \end{cases}
\end{aligned}$$

## Equations


$$ x_{i1} =
  \begin{cases}
                                   1 & \text{if the $i$th person is Asian} \\
                                   0 & \text{if the $i$th person is not Asian}
  \end{cases}$$


$$x_{i2} =
  \begin{cases}
                                   1 & \text{if the $i$th person is Caucasian} \\
                                   0 & \text{if the $i$th person is not Caucasian}
  \end{cases}$$
  
---


$$\begin{aligned}
    y_{i} &= {\beta_{0}} + {\beta_{1}X_{i1}} + {\beta_{2}X_{i2}} + {\epsilon_i} \\
          &=  \begin{cases}
                    {\beta_{0}} + {\beta_{1}} + {\epsilon_i} & \text{if the $i$th person is Asian} \\
                    {\beta_{0}} + {\beta_{2}} + {\epsilon_i} & \text{if the $i$th person is Caucasian}\\
                    {\beta_{0}}  + {\epsilon_i} & \text{if the $i$th person is African American}\\
              \end{cases}
\end{aligned}$$


## Slide with Plot

```{r pressure}
plot(pressure)
```

