---
title: "Classification Methods (LDA)"
author: "Cassidy and Delia"
date: "3/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
require(tidyverse)
require(modelr)
require(MASS)
require(class)
require(stats)
require(knitr)
```

# Read in the data

Start by reading in the midge data as a tibble. 

```{r echo = TRUE}
# read in the dataset
midge.train <- as_tibble(read.csv("Chapter 1 Midge.txt", 
                                  header = TRUE))

```

# Creating the grid and then graphing the midges on to this grid

```{r echo = FALSE}
# create a grid of equally spaced WingLengths and
# AntennaLengths
midge.train %>% 
  data_grid(WingLength = 
              seq_range(WingLength, by = 0.01),
            AntennaLength = 
              seq_range(AntennaLength, by = 0.01)
  ) -> midge.grid
```

```{r echo = FALSE}
midge.train %>% 
  ggplot(aes(x = WingLength, 
             y = AntennaLength)) +
  geom_point(aes(color = Type,
                 shape = Type), size = 2)   -> baseplot

```

```{r echo = FALSE}
baseplot +
  geom_point(data = midge.grid,
             shape = 3, size = 0.5, alpha = 0.3) +
  theme(panel.background = element_blank())
```

Now, we will run LDA.  The `lda()` function comes from the package `MASS`.  In this function, we need to include the formula we want it to use and data we want it to use.  Then, run `summary()`, so we are able to look at the data from our model.

```{r echo = TRUE}
midge_lda <- lda(Type ~ WingLength + AntennaLength, data = midge.train)
summary(midge_lda)

```

Now, we will find the averages of each of each of the categories for each type of midge. 

```{r echo = TRUE}
kable(midge_lda$means, caption = "Group Means")
```

Next, we will find the relevant coefficients.

```{r echo = TRUE}
kable(midge_lda$scaling, caption = "Coefficients of Linear Discriminants")
```

We can also find the prior probabilities of the groups.

```{r echo = TRUE}
kable(midge_lda$prior, caption = "Prior Probabilities of Groups(Pi Hat 1 and Pi Hat 2)")

```


Now, we will make predictions for each point on the grid and color them according to that prediction.  Then, we will draw a boundary line.

```{r echo = TRUE}
predictions_LDA <- predict(midge_lda, midge.grid)
Model_LDA_test_data <-
  cbind(midge.grid,
        .pred = predictions_LDA$class,
        .post = predictions_LDA$posterior,
        .x = predictions_LDA$x[,1])


baseplot +
  geom_point(data = Model_LDA_test_data,
             aes(color = .pred),
             shape = 3, size = 0.5) +
  geom_contour(data = Model_LDA_test_data,
               aes(z = as.numeric(.pred) ),
               breaks = c(1.5),
               color = "black") +
  theme(panel.background = element_blank())

```

