---
title: "ISLR Ch 2 Lab k KNN Classification of Midges"
author: "Daniel Camacho"
date: "February 20, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)     #for ggplot, pipes, etc.
require(modelr)        #for data grid
require(class)         #for knn
```



# Applying KNN Classification to Midge data

## Directions: 

1. Put you name(s) in the YAML header above.

1. Read through and run the code to
understand what it does.

2. Answer the questions at the end, putting your code and answers in this RMarkdown file.



##  Lab Setup:  Function for drawing circles

The function `addCircle()` is used below to illustrate the method of K Nearest Neighbors (KNN) classification, but is not necessary for executing KNN classification.

For the illustrations below, we need to draw a circle from center and point on circumference,
so let's create a function for that.

```{r}
addCircle <- function(center, point) {
  # Calculate the radius between the center and point on circumference
  r <- sqrt((center[[1]] - point[[1]])^2 +
            (center[[2]] - point[[2]])^2)

  # Use parametric function of a circle to create 180 points
  # on the circle centered at "center" of radius "r"
  myCircle <-
    tibble(x = center[[1]] + r*cos(seq(0, 2*pi, by = pi/180)),
           y = center[[2]] + r*sin(seq(0, 2*pi, by = pi/180)))

  # Use geom_path() to draw a path through the points
  # make the circle red and dashed and thin
  geom_path(data = myCircle,
            aes(x = x, y = y),
            color = "red", linetype = "dashed", size = 0.1)
}
```

Here is an example of how to use addCircle
In `ggplot()`, `coord_fixed(`) makes an aspect ratio 1:1
so that circles look like circles instead of ellipses,

```{r}
ggplot() + addCircle(c(2,5), c(7,1)) +
  addCircle(c(0,0), c(10,1)) +
  coord_fixed()
```


## The `midge` data

### Examine the `midge` training data

We will read in the same midge data as for the previous midge lab. Then we will look at the data, in particular, the `Type` variable.

```{r}
midge.train <- as_tibble(read.csv("ISLR Chapter 2 Lab 3 Midge.txt",
                                  header = TRUE))

midge.train 

levels(midge.train$Type)

table(midge.train$Type)
```

Create and save a plot of the midges.

```{r}
midge.train %>%
  ggplot(aes(x = WingLength,
             y = AntennaLength)) +
  geom_point(aes(color = Type,
                 shape = Type), size = 2)   -> midgePlot

midgePlot
```

### Create test data 

Here we will create a set of midge data that we want to classify as either Af or Apf.  The types of this data are unknown, and we will add these midges to the plot.

```{r}
midge.test <- tibble(Type = c("unknown", "unknown", "unknown"),
                     WingLength = c(1.80, 1.84, 2.04),
                     AntennaLength = c(1.24, 1.28,1.4),
                     Index = c(1, 2, 3))
midge.test

midgePlot +
  geom_point(data = midge.test,
             aes(color = Type, shape = Type)) +
  geom_text(data = midge.test,
            aes(label = Index),
            nudge_x = 0.02)
```

### Create grid points

To see how the KNN classification divides midges into Af and Apf, we will create a grid of points to fill the plot above.  The `modelr` command `data_grid` will do this for us by creating a grid as wide and tall as the `WingLength` and `AntennaLength` in `midge.train`.  In this case we are making grid points that are 0.01 mm apart.  Save this into `midge.grid`.  Add the grid to the midge plot and look at it

Think of `midge.grid` as a large set of midges for which the type is unknown.  We will classify all of the midges in this grid and plot the result.

Creating the grid is necessary to see the decision boundary below.

```{r}
midge.train %>%
  data_grid(WingLength =
              seq_range(WingLength, by = 0.01),
            AntennaLength =
              seq_range(AntennaLength, by = 0.01)
  ) -> midge.grid

midgePlot +
  geom_point(data = midge.grid,
             shape = 3, size = 0.5, alpha = 0.3) +
  theme(panel.background = element_blank())
```




 How does KNN work?
 plots circles around the test midge points
 through each of the known midges
 so that we can determine the classification of the
 k nearest neighbors for various k

 Choose which test point to use and store the center
```{r}
j <- 1
center <- c(midge.test[j, 2], midge.test[j,3])
center
```

 Initialize myPlot
```{r}
myPlot <- midgePlot
```

loop through the known midges
adding a circle through the known midge centered
at the test midge
```{r}
for (i in 1:15){
  point <- c(midge.train[i, 2], midge.train[i, 3])
  myPlot <- myPlot + addCircle(center, point)
}
```

Plot the circles, test midge and square up the axes
```{r}
myPlot +
  geom_point(data = slice(midge.test, j)) +
  coord_fixed()
```

### Exercise 

 Repeat with the other 2 test midges


### Using R to do KNN classification

Classify all points in the grid 
`knn()` requires data frames with
no extra columns

Creating the data frame for the training inputs (known midges).
```{r}
train.input <- midge.train %>% select(WingLength, AntennaLength)
```

Creating the data frame for the test inputs (unknown midges = data grid)
```{r}
test.input <- midge.grid %>% select(WingLength, AntennaLength)
```

creating the data frame for the training outputs - must be a vector, not a tibble or data frame. Add the grid to the midge plot and look at it

```{r}
train.output <- midge.train$Type

neighbors <- 3

set.seed(1)
midge.grid$Type <- knn(train = train.input,
                      test  = test.input,
                      cl    = train.output,
                      k     = neighbors,
                      prob = TRUE)


# View(data.frame(attr(midge.grid$Type, 'prob')))

midgePlot +
  geom_point(data = midge.grid,
             aes(color = Type),
             shape = 3, size = 0.5) +
  geom_contour(data = midge.grid,
               aes(z = as.numeric(Type) ),
               breaks = c(1.5),
               color = "black") +
  theme(panel.background = element_blank()) +
  geom_point(data = midge.test,
             color = "black")
```

What other output do we get?
```{r}

for (neighbors in c(3,5,7,9)) 
  {
    set.seed(1)
    midge.grid$Type <- knn(train = train.input,
                          test  = test.input,
                          cl    = train.output,
                          k     = neighbors,
                          prob = TRUE)
    
    
    # View(data.frame(attr(midge.grid$Type, 'prob')))
    
    midgePlot +
      geom_point(data = midge.grid,
                 aes(color = Type),
                 shape = 3, size = 0.5) +
      geom_contour(data = midge.grid,
                   aes(z = as.numeric(Type) ),
                   breaks = c(1.5),
                   color = "black") +
      theme(panel.background = element_blank()) +
      geom_point(data = midge.test,
                 color = "black")
}
```


```{r}
view(data.frame(attr(midge.grid$Type, 'prob')))
```


## Questions



1. What happens with ties in KNN?  Clearly
    explain the options for ties and how
    each works.

2. What happens when we use even values of
     k to classify the midges?  What does
     the decision boundary look like?
     Can we change the behavior of ties to
     change the boundary?  Provide examples
     in R code.

